---
title: "MA388 Sabermetrics: Lesson 19"
subtitle: "Comparing Peak Trajectories - Part II"
author: "LTC Jim Pleuss"
format:
  html:
    theme: cosmo
    toc: true
  pdf:
    documentclass: article
execute:
  echo: true
  warning: false
  message: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# Lesson Objectives

1. Map a quadratic model function over multiple players to compare career trajectories.
2. Identify which historical home run hitter had the highest estimated peak HR rate.
3. Recognize unusual trajectory shapes and hypothesize explanations (e.g., injuries, era effects).

## Comparing Trajectories of Home Run Hitters

1. Find the ten players in baseball history who have hit the most career home runs.

```{r}
library(tidyverse)
library(Lahman)
library(knitr)
```

```{r}
# ***Your code here.***

```

2. Fit quadratic functions to the home run rates of the ten players, where HR.RATE = HR / AB.  Display the fitted trajectories in a single figure.

```{r}
# ***Your code here.***

```

3. On the basis of your work, which player had the highest estimated home run rate at his peak?  Which player among the ten had the smallest peak home run rate?

```{r}
# ***Your code here.***

```

*Your answer here.*

4. Do any of the players have unusual career trajectory shapes?  Is there any possible explanation for these unusual shapes?

*Your answer here.*

