---
title: "MLB Playoff Bracket"
---

::: {=html}
<style>
.pf-wrap { max-width: 1080px; margin: 0 auto; }
.pf-hero { background: linear-gradient(135deg, #12263f 0%, #1d4266 45%, #7b1d2b 100%); color: #fff; border-radius: 14px; padding: 26px 30px; box-shadow: 0 10px 28px rgba(18,38,63,.25); }
.pf-hero h2 { margin: 0; font-size: 1.9rem; font-weight: 700; letter-spacing: .2px; }
.pf-hero p { margin: .45rem 0 0; opacity: .88; font-size: .98rem; }
.pf-grid { display: grid; grid-template-columns: minmax(0, 1.3fr) minmax(0, 1fr); gap: 22px; margin-top: 22px; align-items: start; }
@media (max-width: 880px) { .pf-grid { grid-template-columns: 1fr; } }
.pf-card { background: #fff; border: 1px solid #e3e6ea; border-radius: 12px; padding: 16px 20px 18px; margin-bottom: 20px; box-shadow: 0 1px 3px rgba(16,24,40,.06); }
.pf-card-ds { border-top: 3px solid #2a6fb0; }
.pf-card-lcs { border-top: 3px solid #6b4fa0; }
.pf-card-ws { border-top: 3px solid #d4a017; }
.pf-card-h { display: flex; align-items: center; justify-content: space-between; font-weight: 700; font-size: 1.04rem; margin-bottom: 10px; }
.pf-tag { font-size: .68rem; font-weight: 700; letter-spacing: .07em; text-transform: uppercase; padding: .22rem .55rem; border-radius: 999px; background: #eef2f7; color: #5b6b7f; }
.pf-series { display: grid; grid-template-columns: 88px 1fr; gap: 10px; align-items: center; padding: 8px 0; border-top: 1px solid #f1f3f5; }
.pf-series:first-child { border-top: none; }
.pf-series-l { font-weight: 600; font-size: .88rem; color: #33415c; }
.pf-series-c { display: grid; grid-template-columns: 1fr 128px; gap: 8px; }
.pf-select { width: 100%; padding: .45rem .55rem; border: 1px solid #cfd6df; border-radius: 8px; background: #fff; font-size: .92rem; }
.pf-select:focus { outline: none; border-color: #2a6fb0; box-shadow: 0 0 0 3px rgba(42,111,176,.16); }
.pf-select:disabled { background: #f4f6f8; color: #9aa5b1; border-color: #e4e8ed; }
.pf-name { width: 100%; padding: .5rem .6rem; border: 1px solid #cfd6df; border-radius: 8px; font-size: .98rem; }
.pf-name:focus { outline: none; border-color: #2a6fb0; box-shadow: 0 0 0 3px rgba(42,111,176,.16); }
.pf-actions { display: flex; gap: 10px; flex-wrap: wrap; margin-top: 4px; }
.pf-btn { border: none; border-radius: 8px; padding: .7rem 1.4rem; font-weight: 700; font-size: 1rem; cursor: pointer; }
.pf-btn-primary { background: linear-gradient(180deg, #c8102e, #a30d26); color: #fff; box-shadow: 0 2px 7px rgba(168,13,38,.32); }
.pf-btn-primary:hover { filter: brightness(1.08); }
.pf-btn-ghost { background: #eef1f5; color: #33415c; }
.pf-btn-ghost:hover { background: #e3e8ee; }
.pf-progress-text { font-size: .86rem; font-weight: 600; color: #5b6b7f; margin-bottom: 6px; }
.pf-bar { height: 8px; background: #e9edf2; border-radius: 999px; overflow: hidden; }
.pf-bar > span { display: block; height: 100%; width: 0; background: linear-gradient(90deg, #2a6fb0, #7b1d2b); transition: width .25s ease; }
.pf-bracket-round { margin-top: 12px; padding-left: 11px; border-left: 3px solid #dfe4ea; }
.pf-bracket-h { font-size: .72rem; font-weight: 700; letter-spacing: .07em; text-transform: uppercase; color: #8a97a8; }
.pf-bracket-row { display: flex; justify-content: space-between; gap: 10px; font-size: .9rem; padding: 3px 0; }
.pf-bracket-t { color: #6b7a8d; }
.pf-bracket-v { font-weight: 600; text-align: right; }
.pf-bracket-round:last-child .pf-bracket-v { color: #8a6d0b; }
.pf-table { width: 100%; border-collapse: collapse; font-size: .88rem; }
.pf-table th { text-align: left; font-size: .72rem; letter-spacing: .05em; text-transform: uppercase; color: #8a97a8; padding-bottom: 4px; }
.pf-table td { padding: .34rem 0; border-top: 1px solid #f1f3f5; vertical-align: top; }
.pf-table td:last-child { text-align: right; font-weight: 700; white-space: nowrap; }
.pf-note { font-size: .84rem; color: #6b7a8d; margin: .7rem 0 0; }
.pf-msg { margin-top: 12px; border-radius: 8px; padding: .7rem .9rem; font-weight: 600; font-size: .94rem; }
.pf-msg-ok { background: #eaf7ef; border: 1px solid #b7e3c6; color: #14663a; }
.pf-msg-err { background: #fdecec; border: 1px solid #f5c2c2; color: #9b1c1c; }
.pf-msg-idle { display: none; }
.pf-scroll { overflow-x: auto; margin-top: .4rem; }
.pf-table2 { border-collapse: collapse; width: 100%; font-size: .88rem; }
.pf-table2 th { text-align: left; font-size: .7rem; letter-spacing: .05em; text-transform: uppercase; color: #8a97a8; padding: 0 .6rem .45rem 0; white-space: nowrap; border-bottom: 1px solid #e6eaef; }
.pf-table2 td { padding: .38rem .6rem .38rem 0; border-top: 1px solid #f4f6f8; vertical-align: middle; white-space: nowrap; }
.pf-rank { color: #9aa5b1; font-variant-numeric: tabular-nums; }
.pf-score { font-weight: 700; font-variant-numeric: tabular-nums; }
.pf-who { font-weight: 600; }
.pf-leader { background: linear-gradient(90deg, rgba(212,160,23,.14), rgba(212,160,23,0)); }
.pf-pill { display: inline-block; padding: .1rem .4rem; border-radius: 5px; font-size: .82rem; border: 1px solid transparent; }
.pf-pill + .pf-pill { margin-left: 4px; }
.pf-ok { background: #eaf7ef; border-color: #b7e3c6; color: #14663a; }
.pf-bad { background: #fdecec; border-color: #f5c2c2; color: #9b1c1c; }
.pf-near { background: #fdf5e4; border-color: #f0dcae; color: #8a6d0b; }
.pf-unknown { background: #f2f4f7; border-color: #e4e8ed; color: #9aa5b1; }
.pf-g { font-variant-numeric: tabular-nums; }
.pf-near-key { color: #8a6d0b; font-weight: 600; }
</style>

<div class="pf-wrap">

<div class="pf-hero">
<h2>Your Postseason Picks</h2>
<p>Division Series, then the League Championship Series, then the World Series. Pick a winner <em>and</em> how many games each series will last.</p>
</div>

<div class="pf-grid">

<div>
<form class="pf-card" id="picksForm">
<div class="pf-card-h"><span>Who are you?</span></div>
<input class="pf-name" type="text" id="name" placeholder="First Last" autocomplete="name">
<div class="pf-card pf-card-ds">
<div class="pf-card-h"><span>Division Series</span><span class="pf-tag">Best of 5</span></div>
<div id="ds"></div>
</div>
<div class="pf-card pf-card-lcs">
<div class="pf-card-h"><span>League Championship Series</span><span class="pf-tag">Best of 7</span></div>
<div id="lcs"></div>
</div>
<div class="pf-card pf-card-ws">
<div class="pf-card-h"><span>World Series</span><span class="pf-tag">Best of 7</span></div>
<div id="ws"></div>
</div>
<div class="pf-actions">
<button type="submit" class="pf-btn pf-btn-primary">Submit Picks</button>
<button type="button" id="resetBtn" class="pf-btn pf-btn-ghost">Reset</button>
</div>
<div class="pf-msg pf-msg-idle" id="picksStatus"></div>
</form>
</div>

<div>
<div class="pf-card">
<div class="pf-card-h"><span>Your Bracket</span></div>
<div class="pf-progress-text" id="pfProgress">0 of 7 series complete</div>
<div class="pf-bar"><span id="pfBar"></span></div>
<div id="pfBracket"></div>
</div>
<div class="pf-card">
<div class="pf-card-h"><span>How Scoring Works</span><span class="pf-tag">28 pts max</span></div>
<table class="pf-table">
<thead><tr><th>Component</th><th>Points</th></tr></thead>
<tbody>
<tr><td>Correct winner</td><td>2</td></tr>
<tr><td>Correct series length</td><td>1</td></tr>
<tr><td>Off by exactly one game</td><td>0.5</td></tr>
<tr><td>Correctly called the sweep</td><td>+1</td></tr>
</tbody>
</table>
<p class="pf-note">Series length points do <strong>not</strong> require the right winner: how long a series lasts doesn't depend on who you pick to win it. A sweep is 3&ndash;0 in the Division Series and 4&ndash;0 in the League Championship and World Series.</p>
</div>
</div>

</div>
</div>

<!--
SETUP (instructor only):
1. Paste your Google Apps Script Web App URL into SCRIPT_URL below.
2. Edit SERIES: give each Division Series its two teams and the valid game counts.
The valid counts are the possible series lengths (3-5 for a best-of-5, 4-7 for a best-of-7).
LCS and World Series teams are derived automatically from earlier picks.
3. If you change the scoring, update the How Scoring Works table above and
the SCORING constants at the top of the script.
4. To show everyone's picks on this page, add this handler to your Apps Script
and redeploy the web app as a new version. It returns the picks sheet as CSV,
which score.R downloads every time the site builds:
      function doGet() {
        var values = sheet().getDataRange().getValues();
        var csv = values.map(function (row) {
          return row.map(function (v) {
            if (v === null || v === undefined) return "";
            if (v instanceof Date) {
              return Utilities.formatDate(v, "America/New_York", "M/d/yyyy HH:mm:ss");
            }
            var s = String(v);
            return /[",\n]/.test(s) ? '"' + s.replace(/"/g, '""') + '"' : s;
          }).join(",");
        }).join("\n");
        return ContentService.createTextOutput(csv)
          .setMimeType(ContentService.MimeType.CSV);
      }
5. Put that web app URL in playoffs/sources.txt (or set PLAYOFF_APPS_SCRIPT_URL).
Then each render refreshes everyone's picks; the workflow also refreshes on a
schedule. Without a source the page falls back to the committed picks CSV, so
it still builds.
6. Fill in winners and series lengths in playoffs/results.csv as series finish.
-->
<script>
(function () {
var SCRIPT_URL = "https://script.google.com/macros/s/AKfycbxQqoHdSXiqk2E37KqHzTrlKnBofuClj3uovAD8dbmRKTrfz30wODStR4w9t10G2489pg/exec";

var SCORING = { winner: 2, length: 1, offByOne: 0.5, sweep: 1 };

var SERIES = [
{ key: "alds1", label: "ALDS 1", teams: ["White Sox", "Guardians"], games: [3, 4, 5] },
{ key: "alds2", label: "ALDS 2", teams: ["Yankees", "Rays"], games: [3, 4, 5] },
{ key: "nlds1", label: "NLDS 1", teams: ["Brewers", "Padres"], games: [3, 4, 5] },
{ key: "nlds2", label: "NLDS 2", teams: ["Dodgers", "Phillies/Braves"], games: [3, 4, 5] },
{ key: "alcs", label: "ALCS", games: [4, 5, 6, 7] },
{ key: "nlcs", label: "NLCS", games: [4, 5, 6, 7] },
{ key: "ws", label: "World Series", games: [4, 5, 6, 7] }
];

var EXPECTED_COLUMNS = 2 + SERIES.length * 2;

var ROUNDS = [
{ title: "Division Series", host: "ds", keys: ["alds1", "alds2", "nlds1", "nlds2"], locked: "Pick the Division Series winners", ready: "Select the winner" },
{ title: "Championship Series", host: "lcs", keys: ["alcs", "nlcs"], locked: "Waiting for Division Series picks", ready: "Select the winner" },
{ title: "World Series", host: "ws", keys: ["ws"], locked: "Waiting for LCS picks", ready: "Select the champion" }
];

var byKey = {};
SERIES.forEach(function (s) { byKey[s.key] = s; });
var picks = {};
var games = {};
var form = document.getElementById("picksForm");
var status = document.getElementById("picksStatus");

function option(value, label) {
var o = document.createElement("option");
o.value = value;
o.textContent = label;
return o;
}

function buildRow(s, round) {
var row = document.createElement("div");
row.className = "pf-series";

var label = document.createElement("div");
label.className = "pf-series-l";
label.textContent = s.label;
row.appendChild(label);

var cell = document.createElement("div");
cell.className = "pf-series-c";

var w = document.createElement("select");
w.id = "pick-" + s.key;
w.className = "pf-select";
w.addEventListener("change", rebuild);
picks[s.key] = w;
if (s.teams) {
w.appendChild(option("", round.ready));
s.teams.forEach(function (t) { w.appendChild(option(t, t)); });
} else {
w.appendChild(option("", round.locked));
}
cell.appendChild(w);

var g = document.createElement("select");
g.id = "games-" + s.key;
g.className = "pf-select";
g.addEventListener("change", rebuild);
games[s.key] = g;
g.appendChild(option("", "Games"));
s.games.forEach(function (n, i) {
g.appendChild(option(String(n), "in " + n + (i === 0 ? " (sweep)" : "")));
});
cell.appendChild(g);

row.appendChild(cell);
return row;
}

function setLocked(sel, teams, lockedMsg, readyMsg) {
var keep = sel.value;
sel.innerHTML = "";
if (teams.length !== 2) {
sel.appendChild(option("", lockedMsg));
sel.disabled = true;
return;
}
sel.appendChild(option("", readyMsg));
teams.forEach(function (t) { sel.appendChild(option(t, t)); });
sel.disabled = false;
if (teams.indexOf(keep) >= 0) sel.value = keep;
}

function pair(a, b) {
var x = picks[a].value;
var y = picks[b].value;
return (x && y) ? [x, y] : [];
}

function completed() {
return SERIES.filter(function (s) {
return picks[s.key].value && games[s.key].value;
}).length;
}

function renderSummary() {
var host = document.getElementById("pfBracket");
host.innerHTML = "";
ROUNDS.forEach(function (r) {
var block = document.createElement("div");
block.className = "pf-bracket-round";
var h = document.createElement("div");
h.className = "pf-bracket-h";
h.textContent = r.title;
block.appendChild(h);
r.keys.forEach(function (k) {
var s = byKey[k];
var row = document.createElement("div");
row.className = "pf-bracket-row";
var t = document.createElement("span");
t.className = "pf-bracket-t";
t.textContent = s.label;
var v = document.createElement("span");
v.className = "pf-bracket-v";
var who = picks[k].value || "TBD";
var len = games[k].value ? "  (" + games[k].value + " g)" : "";
v.textContent = who + len;
row.appendChild(t);
row.appendChild(v);
block.appendChild(row);
});
host.appendChild(block);
});

var done = completed();
document.getElementById("pfProgress").textContent = done + " of 7 series complete";
document.getElementById("pfBar").style.width = Math.round((done / 7) * 100) + "%";
}

function rebuild() {
setLocked(picks.alcs, pair("alds1", "alds2"), ROUNDS[1].locked, ROUNDS[1].ready);
setLocked(picks.nlcs, pair("nlds1", "nlds2"), ROUNDS[1].locked, ROUNDS[1].ready);
var finals = (picks.alcs.value && picks.nlcs.value) ? [picks.alcs.value, picks.nlcs.value] : [];
setLocked(picks.ws, finals, ROUNDS[2].locked, ROUNDS[2].ready);
renderSummary();
}

function show(msg, kind) {
status.className = "pf-msg pf-msg-" + kind;
status.textContent = msg;
}

form.addEventListener("submit", function (ev) {
ev.preventDefault();
var name = document.getElementById("name").value.trim();
var missing = [];
if (!name) missing.push("your name");
SERIES.forEach(function (s) {
if (!picks[s.key].value) missing.push(s.label + " winner");
if (!games[s.key].value) missing.push(s.label + " length");
});
if (missing.length) {
show("Still needed: " + missing.join(", ") + ".", "err");
return;
}

var payload = { name: name, submitted_at: new Date().toISOString() };
SERIES.forEach(function (s) {
payload[s.key] = picks[s.key].value;
payload[s.key + "_games"] = games[s.key].value;
});

show("Saving your picks...", "ok");
fetch(SCRIPT_URL, {
method: "POST",
headers: { "Content-Type": "text/plain;charset=utf-8" },
body: JSON.stringify(payload)
})
.then(function (r) { return r.json(); })
.then(function (res) {
if (!res || res.status !== "ok") {
show("The server rejected this: " + ((res && res.message) || "no response") + " Nothing was saved.", "err");
return;
}
if (res.columns !== EXPECTED_COLUMNS) {
show("Saved, but the Apps Script looks out of date: it reported " +
(res.columns === undefined ? "no" : res.columns) + " columns instead of " +
EXPECTED_COLUMNS + ". The games columns may be missing. Redeploy the web app, then submit again.", "err");
return;
}
show("Thanks, " + name + "! Your picks are locked in.", "ok");
form.reset();
rebuild();
})
.catch(function () {
show("Could not reach the Apps Script. Check your connection and try again.", "err");
});
});

document.getElementById("resetBtn").addEventListener("click", function () {
form.reset();
rebuild();
});

ROUNDS.forEach(function (r) {
var host = document.getElementById(r.host);
r.keys.forEach(function (k) { host.appendChild(buildRow(byKey[k], r)); });
});

rebuild();
})();
</script>
:::

<!--
UPDATING THE STANDINGS (instructor only):
1. After a series is played, add the winner and the number of games it took to
playoffs/results.csv, e.g.  alds1,White Sox,4
2. Refresh the picks and rescore by running:
     Rscript playoffs/score.R
   With PLAYOFF_SHEET_ID set it re-downloads the sheet first; without it, it
   re-scores the committed playoffs/data/picks.csv.
3. The tables below are rebuilt on every render (freeze is off for these
   chunks), so the next build shows the updated scores.
-->

## Standings

```{r}
#| label: playoff-standings
#| freeze: false
#| echo: false
#| output: asis
pf_dir <- if (file.exists("score.R")) "." else "playoffs"
if (!file.exists(file.path(pf_dir, "score.R"))) stop("playoffs/score.R not found")
source(file.path(pf_dir, "score.R"))
pf_results <- read_results(file.path(pf_dir, "results.csv"))
pf_picks_path <- file.path(pf_dir, "data", "picks.csv")
pf_picks <- if (file.exists(pf_picks_path)) dedupe_picks(read_picks_csv(pf_picks_path)) else data.frame()
pf_tab <- if (nrow(pf_picks)) score_all(pf_picks, pf_results) else data.frame()
cat(standings_html(pf_tab, pf_results))
```

## Everyone's Picks

```{r}
#| label: playoff-picks-table
#| freeze: false
#| echo: false
#| output: asis
cat(picks_html(pf_picks, pf_tab, pf_results))
```